---
title: "Homework 2"
author: "John DeFilippo"
date: "September 29, 2015"
output: html_document
---

JD: Because Rmarkdown documents are meant to be self contained, all of the 
necessary libraries and data files (if you were using external data files) must
be loaded at the top of the file. When this is done, anyone (provided that they 
have the necessary data and the correct packages installed) will be able to 
render the document on their own computer. 

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)
```

You can also hide code chunks if you would like using, echo = FALSE:

```{r, echo = FALSE}
# This line of code will never show up in the final product.
```


1) Use points and colors to identify continents

```{r}
ggplot(data = gapminder, aes(x = continent, fill = continent)) +
  geom_histogram()

# JD: I think here you were looking for 
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, shape = continent, colour = continent)) + 
  geom_point()

```

2) The per capita gdp has a very large range; use a transformation to linearize the data

```{r}
str(gapminder)

ggplot(data = gapminder, aes(x = log10(gdpPercap), y=lifeExp, colour = continent)) +
  geom_point()

# An alternative is to change the scale of the x axis:
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, colour = continent, shape = continent)) + 
  geom_point() +
  scale_x_log10()
```

3) Include a simple linear fit to the transformed data

```{r}
ggplot(data = gapminder, aes(x = log10(gdpPercap), y=lifeExp, colour = continent, shape = continent)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)

# JD: here you have multiple lines showing up for each continent. There are several
# ways to avoide this. 

# 1 - you can move the colour (which inherently groups the data) aesthetic to 
# refer to just the points:

ggplot(data = gapminder, aes(x = log10(gdpPercap), y = lifeExp)) + 
  geom_point(aes(colour = continent, shape = continent)) +
  geom_smooth(method = 'lm')

# 2 - you can tell the geom_smooth to use only a single group:
ggplot(data = gapminder, aes(x = log10(gdpPercap), y=lifeExp, colour = continent, shape = continent)) +
  geom_point() +
  geom_smooth(method = lm, aes(group = 1))

# If you want more info on the group argument of aes() check out ?ggplot2::group

```

4) Plot the density functions of life expectancy for each continent

```{r}
ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) +
  geom_density(alpha = 0.5) +
  facet_wrap( ~ year, ncol = 4)

# JD: just added some transparency to the density plots to make them prettier.
```

5) Fix the bottom continent labels

```{r}
ggplot(data = gapminder, aes(x = continent, y = lifeExp, colour = continent)) +
  geom_boxplot() +
  facet_wrap( ~ year, ncol = 4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# JD: added some colour to group by continent to make them stand out.
# Nice changing the bottom text. I often prefer an angle of 45 degrees because 
# it can be easier to read.
```

6)  

```{r}
ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) +
  geom_density(alpha = 0.5)

# JD: added transparency to allow us to see the different data layers. 
```

7) Plot the mean life expectancy on a density plot for Asia

```{r}
ggplot(data = gapminder, aes(x = lifeExp, fill = (continent[3]))) +
  geom_density() +
  geom_vline(aes(xintercept = mean(lifeExp)))

# JD:
# This is actually the life expectancy for all continents. Fill is just 
# colouring it the '3rd' colour that ggplot defaults to.
# To get a subset of the data you could do several things, but two options are:
ggplot(data = subset(gapminder, continent == 'Asia'), aes(x = lifeExp)) + 
  geom_density()

# OR using dplyr
ggplot(data = filter(gapminder, continent == 'Asia'), aes(x = lifeExp)) + 
  geom_density

# - There is no need to specify the fill when there is only one group, ggplot 
# will automatically colour it. But, if you wanted to change the colour, you 
# could specify it manually in the geom_density() call.
# e.g., 
geom_density(colour = 'pink')

```
                        
8a) Create a data frame of the mean life expectancies for each continent

```{r}
data.frame(gapminder) %>% 
  group_by(continent) %>%
  summarize(mean_lifeExp = mean(lifeExp))

# JD: There is no need to make gapminder a data.frame because it already is. I 
# think what you are thinking of is when we used as.data.table() in class. This 
# made gapminder into a data.table which is faster (important for bigger datasets 
# or when applying a slow calculation).

```
8b) Plot the density plot of life expectancies for each continent and draw a vertical line to mark the mean life expectancy for each continent.

```{r}
# I can't get the mean to behave independently on each facet. I know it has something to do with creating the data.frame, but I couldn't get it to work.

ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(lifeExp))) +
  facet_wrap( ~ continent, ncol = 3)

# Awesome! You fixed it!
```

#### Find the mistakes

**Note** Here I'm going to use eval = FALSE to prevent the erroneous code from 
running. If I don't do this, R will get mad and the document won't render. But, 
because I have not set echo = FALSE, we can still see the broken code, R just 
won't do anything with it.  

```{r, eval = FALSE}
mean_lifeExp <- mean(hw_gapminder$lifeExpe) # Typo... should be lifeExp
```

```{r}
mean_lifeExp <- mean(hw_gapminder$lifeExp)
```


```{r, eval = FALSE}
# need 2nd concatenate
small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), ('country', 'continent', 'year')]
```

```{r}
small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), c('country', 'continent', 'year')]
```

```{r, eval = FALSE}
mean_gdp <- mean(hw_gapminder$gdpPercap)
```

```{r}
summary(hw_gapminder$gdpPercap)
# has 5 NAs that need to be removed
mean_gdp <- mean(hw_gapminder$gdpPercap, na.rm=TRUE)
```

```{r, eval = FALSE}
max_country <- hw_gapminder$country[which(hw_gapminder$lifeExp = max(hw_gapminder$lifeExp))]
# need == for equivalence
```
```{r}
max_country <- hw_gapminder$country[which(hw_gapminder$lifeExp == max(hw_gapminder$lifeExp))]
# I get Japan, but also 142 levels

# JD: This is because I set some 'chunk options' at the top of my document that 
suppresses messages and warnings for a tidier document.
```